---
title: "Length-based spawning potential ratio (LBSPR)"
output:
  pdf_document:
    extra_dependencies:
      tabularx: null
      booktabs: null
      colortbl: null
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, fig.align = "center", dpi=300,  fig.pos = "H")
```

# Method description

The length-based spawning potential ratio method (LBSPR) allows to estimate the
stock status for data-limited fisheries based only on length frequency data and
some information about the life history parameters. Besides gear selectivity,
the method estimates the fishing mortality relative to natural mortality (F/M)
and the spawning potential ratio (SPR). SPR describes the ratio of the current
spawning stock biomass and the theoretical spawning stock biomass if there was
no fishing. Thus, an SPR of 0% would indicate that there are no mature female
individuals are left. Suggested SPR reference points are a limit at 20% and a
target of 40% (Hordy et al. 2015). The method does not require information about
the uncertain natural mortality directly, but instead uses the ratio of natural
mortality and the von Bertalanffy growth coefficient (K) (M/K). This ratio is
believed to be less variable across stocks and species than the individual
parameters and thus assumed to be 1.5 (Prince et al. 2015). However, other
studies suggest that this ratio is not invariant but varies systematically based
on the timing of maturation (e.g. Thorson et al. 2017).

LBSPR makes the following assumptions:

- Representative length measurements: The method assumes that the dataset is
  representative of the length distributions of the whole catch. This means that
  either the length of all individuals in the catch were measured or a
  randomised subsample of the catch was measured.

- Equilibrium conditions: The routine assumes constant recruitment and natural
  mortality as well as somatic growth and maturation over time, i.e. within the
  year and over all years covered by the data set.

- Density dependence: The routine assumes density independent maturity and
  somatic growth.

- Length-independent natural mortality: The method assumes that the natural
  mortality is equal for all length classes.

- Somatic growth follows VBG: The method assumes that the growth of individuals
  in length follows the logistic von Bertlanffy growth (VBG) function. This is
  an often made assumption for the growth of fish, but might not reflect well
  the growth of species with an exoskeleton, such as crustaceans, nor the growth
  of early life stages of fish.

- Closed population: The routine assumes that the stock (population) under study
  is closed, meaning that there is no immigration or emigration taking place.
  Immigration and emigration can both bias estimated mortality rates and stock
  status. Furthermore, fish migrations often vary for various life stages and
  might thus affect the representativeness of the length measurements if the
  population is not closed.

&nbsp;
&nbsp;
&nbsp;

# Summary of input data and parameters

The uploaded data (*"*`r paste0("*",input$fileLBSPR$name,"*")`*"*) spans `r length(lbspr_dat$dataExplo[['lfq']]$dates)` year(s). Following years were selected for the analysis: `r paste(input$LBSPR_years_selected, collapse = ", ")`. Figure 1 shows the length frequency distributions for each year that were binned with a bin size of `r input$LBSPR_binSize` `r input$LBSPR_lengthUnit`.

&nbsp;
&nbsp;
&nbsp;

```{r echo=FALSE, fig.width=9, fig.height=7}
plotLBSPR.data(lbspr_dat, input)
```

```{r echo=FALSE,results='asis'}
writeLines(captionLBSPR.plots(lbspr_dat, input,
                              format = "withFigLatex", type = "data"))
```



&nbsp;
&nbsp;
&nbsp;

The life-history parameters required for LBSPR were set to the values presented
in Table 1.


```{r echo=FALSE}
tableLBSPR.inputPars(lbspr_dat, input, format = "kable")
```

&nbsp;
&nbsp;
&nbsp;


# Results

The estimated SPR, selectivity parameters (L~s50~ and L~s95~), and the ratio of
fishing to natural mortality rate (F/M) are provided in Table 2. The estimated
95% confidence interval for each parameter is provided in brackets. If the
uploaded data spans several years, the table contains multiple rows and each row
contains the estimates for the specific year.

&nbsp;
&nbsp;
&nbsp;

```{r echo=FALSE}
tableLBSPR.results(lbspr_dat, input, format = "kable")
```

&nbsp;
&nbsp;
&nbsp;

Figure 2 shows the estimated spawining potential ratio (SPR) value in relation
to the SPR reference points. The estimated SPR is indicated by the dashed line
with the SPR value next to the dashed line. The colored areas indicate the SPR
reference points: The red area/line indicates the proportion of SPR below the
limit reference point; the green area/line indicates the proportion above the
limit and below the target reference point; and the the yellow area indicates
the proportion above the target reference point. Note, that when the assessment
is done for multiple years, only the last year of the assessment is shown in
this graph.

&nbsp;
&nbsp;
&nbsp;


```{r echo=FALSE, fig.width=5, fig.height=5}
plotLBSPR.pie(lbspr_dat, input)
```

```{r echo=FALSE,results='asis'}
writeLines(captionLBSPR.plots(lbspr_dat, input,
                              format = "withFigLatex", type = "pie"))
```

&nbsp;
&nbsp;
&nbsp;

Figure 3 presents the provided maturity information (black line), as well as the
estimated selectivity information (colored lines). The curves indicate the
proportion of the stock that is mature or vulnerable to the gear (y axis) at a
given length (x axis).

&nbsp;
&nbsp;
&nbsp;


```{r echo=FALSE, fig.width=5, fig.height=4}
plotLBSPR.sel(lbspr_dat, input)
```

```{r echo=FALSE,results='asis'}
writeLines(captionLBSPR.plots(lbspr_dat, input,
                              format = "withFigLatex", type = "sel"))
```

&nbsp;
&nbsp;
&nbsp;

Figure 4 shows a time series plot of the estimated selectivity parameters,
fishing mortality relative to natural mortality, and spawning potential ratio
(SPR). The lines correspond to the smoothed estimates over time. Note, that the
graphs only show single points when the assessment spans a single year.

&nbsp;
&nbsp;
&nbsp;

```{r echo=FALSE, fig.width=9, fig.height=5}
plotLBSPR.ts(lbspr_dat, input)
```

```{r echo=FALSE,results='asis'}
writeLines(captionLBSPR.plots(lbspr_dat, input,
                              format = "withFigLatex", type = "ts"))
```

&nbsp;
&nbsp;
&nbsp;



# References

Hordyk, A., Ono, K., Valencia, S., Loneragan, N., & Prince, J. (2015). A novel
length-based empirical estimation method of spawning potential ratio (SPR), and
tests of its performance, for small-scale, data-poor fisheries. ICES Journal of
Marine Science, 72(1), 217-231. https://doi.org/10.1093/icesjms/fsu004

Hordyk, A., Ono, K., Sainsbury, K., Loneragan, N., & Prince, J. (2015). Some
explorations of the life history ratios to describe length composition,
spawning-per-recruit, and the spawning potential ratio. ICES Journal of Marine
Science, 72(1), 204-216. https://doi.org/10.1093/icesjms/fst235

ICES (2018). ICES Technical guidance for providing reference points for stocks
in categories 3 and 4. ICES Technical Guidelines.
https://doi.org/10.17895/ices.pub.4128

Prince, J., Hordyk, A., Valencia, S. R., Loneragan, N., & Sainsbury, K. (2015).
Revisiting the concept of Beverton–Holt life-history invariants with the aim of
informing data-poor fisheries assessment. ICES Journal of Marine Science, 72(1),
194-203. https://doi.org/10.1093/icesjms/fsu011

Thorson, J. T., Munch, S. B., Cope, J. M., & Gao, J. (2017). Predicting life
history parameters for all fishes worldwide. Ecological Applications, 27(8),
2262-2276. https://doi.org/10.1002/eap.1606
