---
title: "Length-based stock assessment with TropFishR"
output:
  pdf_document:
    extra_dependencies:
      tabularx: null
      booktabs: null
      colortbl: null
date: "`r format(Sys.time(), '%d %B, %Y')`"
keep_tex: true
header-includes:
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, fig.align = "center", fig.pos = "H")
```


# Method description

The length-based stock assessment with TropFishR is based on the routine
outlined in the Technical report Introduction to Tropical Fish Stock Assessment
(FAO, 1998), and compiled into the R package TropFishR by Mildenberger et al.
(2017). The assessment routine consists of 4 consecutive steps and methods:

## Estimation of the growth parameters

The study of fish growth involves a determination of body size as a function of
age. Most stock assessment methods work essentially with age composition data.
In temperate waters it is easier to acquire data to estimate age by counting
year rings on hard parts of organisms such as scales and otoliths (ear bones).
These rings are formed due to strong seasonal fluctuations in environmental
conditions. In tropical areas such drastic changes do not occur and it is
therefore very difficult, if not impossible to use these kind of seasonal rings
for age determination. The Electronic LEngth Frequency ANalysis (ELEFAN) is one
of a number of numerical methods that have been developed, which allow the
conversion of length-frequency data into age composition. Although these methods
do not require the reading of rings on hard parts, the final interpretation of
the results becomes much more reliable if at least some direct age readings are
available. ELEFAN fits a seasonal version of the von Bertalanffy Growth Function
(VBGF) by:

- restructuring the length-frequency data using a procedure that scores the
  length bins based on their deviation from a moving average (MA) across
  neighboring bins.

- calculating a score value as the cumulative score for a given set of VBGF
  parameters based on the sum of the individual bin scores that are intersected
  by resulting growth curves.

- optimising over VBGF parameters maximising the score value by a genetic
  algorithm (GA). Imitating the natural process of survival of the fittest, the
  GA simulates a population of random combinations of growth parameters in which
  only the individuals (or combinations) with the highest fitness value (score
  value; Rn) can pass on their parameters to the next generations. The
  combination of growth parameters with the highest score value after a set
  number of generations is defined as the best estimate. For more information
  about the genetic algorithm please refer to Scrucca (2013).

## Estimation of the natural mortality rate

The natural mortality rate (M) is estimated by an empirical formula based on
estimated growth parameters. The options are:

- Then's growth formula (Then et al. 2015),
- Pauly's growth and temperature formula (Pauly 1980), and
- Then's maximum age formula (Then et al. 2015);
- Gislason's length-based formula (Gislason et al. 2010);
- Lorenzen's length-based formula (Lorenzen et al. 2022);

While the first and the last two options do not require any additional
information, the second requires the average annual sea surface temperature
(SST) in degrees Celsius and allows for a correction for schooling fish
(multiplication with 0.8). The third option requires an estimate of the maximum
age of the fish.

## Estimation of total and fishing mortality rates

The length-converted catch curve is used to estimate the total mortality rate
(Z) and the fishing mortality rate (F) based on M: F = Z - M.

## Estimation of the stock status

The Thompson and Bell yield per recruit (YPR) model is used to estimate yield
and biomass per recruit as well as the spawning potential ratio (SPR) over a
range of fishing mortality rates and selectivity definitions. The stock status
is based on current F relative to biological reference points, such as
\(F_{max}\), \(F_{0.1}\), \(F_{35}\). The estimation of SPR requires information
about the maturity parameters.


## Assumptions of TropFishR

The above mentioned methods make specific assumptions about the data, the stock,
or the fisheries targeting the stock. It is important to be aware of these
assumptions and, thus, the limitations of the results.

- Representative length measurements: The routine assumes that the dataset is
  representative of the length distributions of the whole catch. This means that
  either the length of all individuals in the catch were measured or a
  randomised subsample of the catch was measured.

- Gear/fleet selectivity: The routine assumes sigmoidal selectivity as it is
  commonly assumed for 'trawl-like' fishing gear. While the selectivity of
  gillnets and hook-based methods might correspond to a bell-shaped curve, the
  combination of various mesh and hook sizes might lead to an overall
  'trawl-like' fleet selectivity.

- Equilibrium conditions: The routine assumes constant recruitment, fishing and
  natural mortality as well as somatic growth and maturation over time, i.e.
  within the year and over all years covered by the dataset.

- Density dependence: The routine assumes density independent maturity and
  somatic growth. All reference points are based on the 'per recruit' model,
  thus no assumptions are made about the stock recruitment relationship.

- Correlation of life-history parameters: The use of empirical formulae for the
  estimation of natural mortality assumes that the growth parameters
  (\(L_\infty\) and K) are reliable predictors of the natural mortality.

- Somatic growth follows VBG: The estimation of growth parameters with ELEFAN
  assumes that the growth of individuals in length follows the logistic von
  Bertlanffy growth (VBG) function. This is an often made assumption for the
  growth of fish, but might not reflect well the growth of species with an
  exoskeleton, such as crustaceans, nor the growth of early life stages of fish.

- Closed population: The routine assumes that the stock (population) under study
  is closed, meaning that there is no immigration or emigration taking place.
  Immigration and emigration can both bias estimated mortality rates and stock
  status. Furthermore, fish migrations often vary for various life stages and
  might thus affect the representativeness of the length measurements if the
  population is not closed.

&nbsp;
&nbsp;
&nbsp;

# Input parameters

The uploaded length data set (\verb|`r input$fileGa$name`|) spans over
`r years = unique(as.numeric(as.character(format(elefan_ga$dataExplo$lfq$dates,"%Y")))); if(length(years) == 1){ paste0("1 year (", years, ")") }else{ paste0(length(years), "years (",paste(range(years), collapse = " - "),")")}` (Fig 1.).

Figure 1 shows the length-frequency distributions per sampling time (x axis),
where panel A in shows the raw data, while panel B shows the restructured data,
i.e. the data after subtracting the moving average (MA) from each length class.
The combination of bin size and MA critically affects the separation of peaks
(i.e. potential cohorts) in the dataset and thus the estimation of growth
parameters by ELEFAN. Blue shading indicates a high count per length bin (panel
A) and a high positive value (panel B). Red shading indicates a negative value
(only panel B). A good bin size value reduces noise in the data by aggregation
and should be defined before the MA value. A good MA value leads to visually
distinct peaks, particularly among the smaller length classes.

```{r echo=FALSE, fig.align="center", dpi=300, fig.width=9, fig.height=8}
plotTropFishR.data(elefan_ga, input)
```
```{r echo=FALSE,results='asis'}
writeLines(captionTropFishR.plots(elefan_ga, input, format = "withFigLatex", type = "explo"))
```

An overview over the input parameters is provided in Table 1.

```{r echo=FALSE}
tableTropFishR.input(elefan_ga, input, format = "kable")
```

&nbsp;
&nbsp;
&nbsp;

# Results

## Growth

Figure 2 shows the same length frequency distributions as Figure 1 overlaid with
the growth curves estimated by ELEFAN. This plot allows the visual inspection of
how well the estimated curves connect the peaks in the length-frequency data
(i.e. potential cohorts) in the raw (A) and restructured (B) data set.

```{r echo=FALSE, fig.align="center", dpi=300, fig.width=9, fig.height=8}
plotTropFishR.growth(elefan_ga, input)
```
```{r echo=FALSE,results='asis'}
writeLines(captionTropFishR.plots(elefan_ga, input, format = "withFigLatex", type = "growth"))
```

Table 2 includes the estimated von Bertlanffy growth parameters \(L_\infty, K,
t_a\) the growth performance coefficient phi', and the best score value (Rn).
The growth performance index is calculated based on the formula \(phi' =
\log_{10} K + 2 \log_{10} L_\infty\) and provides a metric that accounts for
the correlation of \(L_\infty\) and \(K\) to compare growth parameters among
analyses or species. The Rn value is the sum of all values associated with the
bins after restructuring that intersect the growth curves. The value can be used
to compare the fit of estimated growth parameters to the data set for different
MA values, growth paramater search spaces, or ELEFAN optimisation parameters
(e.g. population size). However, it cannot be used to compare the fit for
different data sets or different bin sizes as this affects the maximum overall
possible Rn value.


```{r echo=FALSE}
tableTropFishR.growth(elefan_ga, input, format = "kable")
```

Figure 3 shows the improvement of the ELEFAN fit in terms of the average and
best score value (fitness value) of the genetic algorithm used in ELEFAN_GA over
the number of iterations (generations). Ideally, the number of iterations
(generations) is large enough so that there are no large jumps visible during
the last iterations of the best and average score value.

```{r echo=FALSE, fig.align="center", dpi=75, fig.width=6, fig.height=5}
plotTropFishR.ga(elefan_ga, input)
```
```{r echo=FALSE,results='asis'}
writeLines(captionTropFishR.plots(elefan_ga, input, format = "withFigLatex", type = "ga"))
```


&nbsp;

## Mortality rates

Figure 4 shows the estimated total, fishing, and natural mortality rates by
length. The dashed lines correspond to the average mortality rates for the main
exploited length classes, i.e. the length classes that are considered in the
regression analysis for the length-converted catch curve. Note, that the natural
mortality rates estimated by the empirical formulae after Then et al. (2015) and
Pauly et al. (1980) are independent of length. By contrast, the formulae after
Gislason et al. (2010) and Lorenzen et al. (2022) estimate length-dependent
natural mortality rates based on the asymptotic length, growth coefficient, and
body length of the fish.

```{r echo=FALSE, fig.align="center", dpi=75, fig.width=6, fig.height=5}
plotTropFishR.mort(elefan_ga, input)
```
```{r echo=FALSE,results='asis'}
writeLines(captionTropFishR.plots(elefan_ga, input, format = "withFigLatex", type = "mort"))
```

Table 3 lists the estimated mortality rates (Z, F, M), the exploitation rate
(E), and the estimated/provided selectivity parameters (\(L_{s50}, L_{s75}\)).

```{r echo=FALSE}
tableTropFishR.mort(elefan_ga, input, format = "kable")
```

Figure 5 shows the logarithm of the catch per length interval against the
relative age (x axis). Highlighted points were used in the regression analysis
of the catch curve for the estimation of the total mortality rate (Z), which
corresponds to the slope of the displayed regression line. The selection of
points is automatic and based on a list of expert recommendations.


```{r echo=FALSE, fig.align="center", dpi=75, fig.width=6, fig.height=5}
plotTropFishR.catchcurve(elefan_ga, input)
```
```{r echo=FALSE,results='asis'}
writeLines(captionTropFishR.plots(elefan_ga, input, format = "withFigLatex", type = "cc"))
```

Figure 6 shows the probability of capture as a measure of the selectivity used
in the assessment. The curve is either based on selectivity parameters provided
by the user (\(L_{s50}\) and \(L_{s75}\) or selection width) or was estimated by the
catch curve. The selection ogive that is displayed is used for the Yield Per
Recruit analysis (YPR).

```{r echo=FALSE, fig.align="center", dpi=75, fig.width=6, fig.height=5}
plotTropFishR.sel(elefan_ga, input)
```
```{r echo=FALSE,results='asis'}
writeLines(captionTropFishR.plots(elefan_ga, input, format = "withFigLatex", type = "sel"))
```




&nbsp;

## Reference points

Table 4 includes the estimated reference points (\(F_{max}, F_{0.1}, F_{0.5}\)),
and the SPR-based reference points (\(F_{30}, F_{35}, F_{40}\)) if the maturity
parameters (\(L_{m50}\) and \(L_{m75}\)) are provided.

```{r echo=FALSE}
tableTropFishR.refs(elefan_ga, input, format = "kable")
```

Figure 7 shows the results of the Yield Per Recruit model as yield and biomass
per recruit in panel A and B, respectively, against a range of fishing mortality
rates (x axis). Grey segments display the estimated reference points. Fmax is
defined as the fishing mortality (F) leading to the maximum yield per recruit.
F0.1 corresponds to F where the slope of the yield per recruit curve is equal to
10% of the slope in the origin and represents a more conservative reference
point than \(F_{max}\). \(F_{0.5}\) corresponds to F where the biomass per
recruit is equal to 50% of the biomass per recruit without fishing. If the
maturity parameters \(L_{m50}\) and \(L_{m75}\) are provided, an optional third
panel (C) shows the Spawning Potential Ratio (SPR) for a range of fishing
mortality rates. Furthermore, the three reference points \(F_{30}, F_{35}, and
F_{40}\)are shown, which correspond to F that leads to an SPR of 30%, 35%, and
40% respectively.

```{r echo=FALSE, fig.align="center", dpi=300, fig.width=9, fig.height=10}
plotTropFishR.ypr(elefan_ga, input)
```
```{r echo=FALSE,results='asis'}
writeLines(captionTropFishR.plots(elefan_ga, input, format = "withFigLatex", type = "ypr"))
```




&nbsp;

## Stock status and management

Table 5 presents the estimated stock status in terms of the current fishing
mortality (F) relative to reference points (\(F_{max}, F_{0.1}, F_{0.5}\)). If
the maturity parameters (\(L_{m50}\) and \(L_{m75}\)) are provided, additional
reference points (\(F_{30}, F_{35}, F_{40}\)) and the current Spawning Potential
Ratio (SPR) are included as well.

```{r echo=FALSE}
tableTropFishR.status(elefan_ga, input, format = "kable")
```

Figure 8 shows the yield per recruit (A) and biomass per recruit (B) for various
fishing mortality rates (x axis) and gear selectivities (y axis). The color
indicates high (red) to low (blue) yield and biomass for the various
combinations of F and L50 values. The black dot indicates the current yield and
biomass.

```{r echo=FALSE, fig.align="center", dpi=300, fig.width=9, fig.height=11}
plotTropFishR.iso(elefan_ga, input)
```
```{r echo=FALSE,results='asis'}
writeLines(captionTropFishR.plots(elefan_ga, input, format = "withFigLatex", type = "iso"))
```



&nbsp;
&nbsp;
&nbsp;


# References

Gislason, H., Daan, N., Rice, J.C. and Pope, J.G., 2010. Size, growth,
temperature and the natural mortality of marine fish. Fish and Fisheries, 11(2),
pp.149-158. https://doi.org/10.1111/j.1467-2979.2009.00350.x

Lorenzen, K., Camp, E. V., and Garlock, T. M., 2022. Natural mortality and body
size in fish populations. Fisheries Research, 252, 106327.
https://doi.org/10.1016/j.fishres.2022.106327

Mildenberger TK, Taylor MH, Wolff M, 2017. TropFishR: an R package for
fisheries analysis with length-frequency data. Methods in Ecology and Evolution,
8:11, 1520-1527. https://doi.org/10.1111/2041-210X.12791

Pauly, D. 1980. On the interrelationships between natural mortality, growth
parameters, and mean environmental temperature in 175 fish stocks. J. Cons. Int.
Explor. Mer: 175-192.

Pauly, D. and N. David, 1981. ELEFAN I, a BASIC program for the objective
extraction of growth parameters from length-frequency data. Meeresforschung,
28(4):205-211

Pauly, D., 1983. Length-converted catch curves: a powerful tool for fisheries
research in the tropics (part I). ICLARM Fishbyte, 1(2), 9-13

Sparre, P. and Venema, S.C., 1998. Introduction to tropical fish stock assessment.
Part 1. Manual. FAO Fisheries Technical Paper, (306.1, Rev. 2): 407 p.

Taylor MH, Mildenberger TK, 2017. Extending electronic length frequency
analysis in R. Fisheries Management and Ecology, 24:4, 230-238.
https://doi.org/10.1111/fme.12232

Then, A. Y., J. M. Hoenig, N. G. Hall, D. A. Hewitt. 2015. Evaluating the
predictive performance of empirical estimators of natural mortality rate using
information on over 200 fish species. ICES J. Mar. Sci. 72: 82-92.

Thompson, W.F. and Bell, H. 1934. Biological statistics of the Pacific halibut
fishery. 2.Effect of changes in intensity upon total yield, and yield per unit
gear. Rep. Internat. Fish. Comm. 8, 48 pp.
