---
title: "Length-based indicators (LBIs)"
output:
  pdf_document:
    extra_dependencies:
      tabularx: null
      booktabs: null
      colortbl: null
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE, fig.align = "center", dpi=300,  fig.pos = "H")
```


# Method description

LBIs rely on the relationships between life history theory and length frequency
data from catches and are based on three simple ideas: (1) the length frequency
data reflects the conservation of large, mature individuals with arguably a high
fecundity (also referred to as 'mega spawners'); (2) the data demonstrates the
conservation of small immature individuals; (3) the length frequency data
consists mainly of fish of the size at which the highest yield from a cohort or
the maximum sustainable yield (MSY) is to be expected. In theory, these metrics
are, thus, indicating how a stock is performing in terms of yield optimisation
and conservation goals (i.e. avoiding growth and recruitment overfishing). The
seven LBIs used for this stock assessment are described in Table 1. The table
also lists the reference point, indicator ratio, expected value and property for
each indicator. Note that the indicator L~mean~ is listed twice as it can be
compared to two different reference points: L~opt~ and L~F=M~. These indicator
ratios and expected values for each indicator are used for the stock status
classification in Table 3 and Figure 2. The indicators have different properties
(last column in Table 1), indicating either whether the stock is likely to show
signs of depletion regarding larger or immature individuals, or whether the
stock is likely to be harvested optimally (optimal yield and MSY).

&nbsp;
&nbsp;
&nbsp;

```{r echo=FALSE}
tableLBI.intro(lbi_dat, input, format = "kable")
```

&nbsp;
&nbsp;
&nbsp;

The indicators make a number of assumptions:

- Representative length measurements: The method assumes that the dataset is
  representative of the length distributions of the whole catch. This means that
  either the length of all individuals in the catch were measured or a
  randomised subsample of the catch was measured.

- Equilibrium conditions: The routine assumes constant recruitment and natural
  mortality as well as somatic growth and maturation over time, i.e. within the
  year and over all years covered by the data set.

- Density dependence: The routine assumes density independent maturity and
  somatic growth.

- Length-independent natural mortality: The method assumes that the natural
  mortality is equal for all length classes.

- Somatic growth follows VBG: The method assumes that the growth of individuals
  in length follows the logistic von Bertlanffy growth (VBG) function. This is
  an often made assumption for the growth of fish, but might not reflect well
  the growth of species with an exoskeleton, such as crustaceans, nor the growth
  of early life stages of fish.

- Closed population: The routine assumes that the stock (population) under study
  is closed, meaning that there is no immigration or emigration taking place.
  Immigration and emigration can both bias estimated mortality rates and stock
  status. Furthermore, fish migrations often vary for various life stages and
  might thus affect the representativeness of the length measurements if the
  population is not closed.

&nbsp;
&nbsp;
&nbsp;


# Summary of input data and parameters

The uploaded data (*"*`r paste0("*",input$fileLBI$name,"*")`*"*) spans `r length(lbi_dat$dataExplo[['lfq']]$dates)` year(s). Following years were selected for the analysis: `r paste(input$LBI_years_selected, collapse = ", ")`. Figure 1 shows the length frequency distributions for each year that were binned with a bin size of `r input$LBI_binSize` `r input$LBI_lengthUnit`.

&nbsp;
&nbsp;
&nbsp;

```{r echo=FALSE, fig.align="center", dpi=300, fig.width=9, fig.height=7}
plotLBI.data(lbi_dat, input)
```

```{r echo=FALSE,results='asis'}
writeLines(captionLBI.plots(lbi_dat, input, format = "withFigLatex", type = "data"))
```

&nbsp;
&nbsp;
&nbsp;

The life-history parameters required for the estimation of the length-based
indicators (LBIs) were set to following values (Table 2).


```{r echo=FALSE}
tableLBI.inputPars(lbi_dat, input, format = "kable")
```


&nbsp;
&nbsp;
&nbsp;


# Results

The estimated LBIs are provided in Table 2. If the uploaded data spans several
years, the table contains multiple rows and each row contains the estimates for
the specific year. Note, that the length at maximum biomass (L~maxy~) is only
estimated if the parameters of the length-weight relationship are provided.
These parameters provide the basis for the indicator ratios in Figure 2 and
Table 3 that indicate the stock status.

&nbsp;
&nbsp;
&nbsp;

```{r echo=FALSE}
tableLBI.indicators(lbi_dat, input, format = "kable")
```


&nbsp;
&nbsp;
&nbsp;

Table 3 contains all estimated LBI ratios indicating the stock status for each
year. The color background of each cell indicates whether there is an indication
for overexploitation (red) or not (green) based on the specific reference points
for each indicator ratio (see Table 1 for more details). Ideally, all cells are
green, thus, showing no indication of overfishing. Note, that green and red
cells in the same year do not mean that the indicators contradict each other,
but could indicate that the stock performs differently regarding different
properties (see Table 1 for more details on the properties of each indicator).
For example, the indicators could indicate that the proportion of older larger
individuals is smaller than expected (red L~95%~/L~inf~), but at the same time
the proportion of immature individuals is as expected (green L~25%~/L~mat~).
Generally, the more indicators are red, the larger the likeliness of
overexploitation.

&nbsp;
&nbsp;
&nbsp;

```{r echo=FALSE}
tableLBI.ratios(lbi_dat, input, format = "kable")
```

&nbsp;
&nbsp;
&nbsp;

Figure 2 is a graphical representation of the indicator ratios over time (if the
uploaded data sets spans multiple years). The reference point is represented by
a horizontal dashed line. Ideally, the indicator ratios are equal or close to
the expected values, e.g. Lmean/Lopt is equal or close to 1. The further the
points/lines are from the expected values of the various indicator ratios, the
more evidence of overfishing. Whenever the graph includes a greenish area the
horizontal line represents a limit reference point, rather than a target
reference point and any value above the limit reference point (i.e. in the
greenish area) is considered desirable.


&nbsp;
&nbsp;
&nbsp;

```{r echo=FALSE, fig.align="center", dpi=300, fig.width=7, fig.height=7}
plotLBI.fit(lbi_dat, input)
```

```{r echo=FALSE,results='asis'}
writeLines(captionLBI.plots(lbi_dat, input, format = "withFigLatex", type = "fit"))
```


&nbsp;
&nbsp;
&nbsp;

# References

Cope, J. M., & Punt, A. E. (2009). Length-based reference points for
data-limited situations: applications and restrictions. Marine and Coastal
Fisheries: Dynamics, Management, and Ecosystem Science, 1(1), 169-186.
https://doi.org/10.1577/C08-025.1

Froese, R. (2004). Keep it simple: three indicators to deal with overfishing.
Fish and fisheries, 5(1), 86-91.
https://doi.org/10.1111/j.1467-2979.2004.00144.x

ICES (2018). ICES Technical guidance for providing reference points for stocks
in categories 3 and 4. ICES Technical Guidelines.
https://doi.org/10.17895/ices.pub.4128

Kell, L. T., Minto, C., & Gerritsen, H. D. (2022). Evaluation of the skill of
length-based indicators to identify stock status and trends. ICES Journal of
Marine Science. https://doi.org/10.1093/icesjms/fsac043
